---
title: "Homework 1"
author: "jiaying Ning"
date: "9/23/2020"
output: html_document
---


**Problem 1**

load packages
```{r, echo = F, message = FALSE, warning=FALSE}
rm(list = ls())
library(arsenal)
library(dplyr)
library(ggplot2)
```

import data
```{r, echo = F, message = FALSE, warning=FALSE}
setwd("/Users/jiayingning/Documents/Columbia\ Fall2020/P8130/Jiaying\ Ning/P8130_jn2761")
Antibodies<-read.csv("Antibodies.csv")
```

1) Using the entire sample, provide descriptive statistics for all variables of interest.
a) Total sample size for each variable (N); Mean/SD, Median/IQR, Min and Max for continuous
variables; Frequency/Percentages for categorical variables; Number of missing values for each
variable. (8p)
b) Generate a his
togram for the Ig-M values and comment on its shape. (2p)


since the only continuous variable is Antibody_Igm. We will be taking the descriptive statistics measure for this continuous variable.

below are its overall summary and desctiptive value table
```{r Problem 1}
#overall summary
summary(Antibodies)

#print out the class table for each variable in the dataset
t(t(sapply(Antibodies, class)))

#print out the descriptive statistics
descriptivestats = summary(na.omit(Antibodies$Antibody_IgM))
descriptivestats
IQR(na.omit(Antibodies$Antibody_IgM))
sd(na.omit(Antibodies$Antibody_IgM))

```

From the dataset, I observe that only meaningful variable that is numeric is "Antibody_IgM". Variable "AgeCategory","Smell","Gender" are all factor.

For continuous variable:
Antibody_Igm has means of `r mean(na.omit(Antibodies$Antibody_IgM))` with standard deviation of `r sd(na.omit(Antibodies$Antibody_IgM))`, median of `r median(na.omit(Antibodies$Antibody_IgM))` and IQR of `r IQR(na.omit(Antibodies$Antibody_IgM))`.Its min is `r min(na.omit(Antibodies$Antibody_IgM))` and max is `r max(na.omit(Antibodies$Antibody_IgM))`

For categorical variable
the frequency and proportion of agegroup is:
```{r freq age}
#frequency for categorical variable
##frequency of age category
table(Antibodies$AgeCategory)
prop.table(table(Antibodies$AgeCategory))
```
the frequency and proportion of subject's smell status is:
```{r freq smell}
##frequency of smell
table(Antibodies$Smell)
prop.table(table(Antibodies$Smell))
```
the frequency and proportion of smell is:
```{r freq gender}
##frequency of Gender
table(Antibodies$Gender)
prop.table(table(Antibodies$Gender))
```

From the summary table we know that there are 1224 NA values in Antibody_Igm variable. 34 unanswered/other in smell category.

```{r histogram IGM}
ggplot(data=Antibodies,mapping=aes((Antibody_IgM))) + geom_histogram()
```
From the graph showing above, we can see that the distribution is skewing to the right. Most people have low antibody_igm. 

**problem 2**

2) Provide descriptive statistics for all variables, stratified by smell category.
a) Provide descriptive statistics for all the other variables by the two smell categories (normal vs
altered) and present them in a tabular form (see example below). Briefly comment on the
differences observed between the two groups. (6p)
Hint: tableby() in R can easily generate this, but feel free to create your own table in Word.

```{r table}
library(dplyr)
library(furniture)

table1(Antibodies,AgeCategory,Subject,Antibody_IgM,Gender,splitby = "Smell",   row_wise = TRUE)

Antibodies_clear = filter(Antibodies, Smell != "Unanswered/Others")
table1(Antibodies_clear,AgeCategory,Subject,Antibody_IgM,Gender,splitby = "Smell",   row_wise = TRUE)


```
Overall, more people reported having their smell altered than not. This difference in proportion is true among all age group and gender. 


b) Use the Ig-M variable to generate side-by-side histograms and boxplots by smell categories
```{r side by side histogram and boxplot}
ggplot(Antibodies_clear,aes(x=Antibody_IgM))+geom_histogram()+facet_grid(~Smell)+theme_bw()

ggplot(Antibodies_clear, aes(x=Smell, y=Antibody_IgM)) + 
    geom_boxplot()

```



(normal vs altered). Make sure you label your figures appropriately and briefly discuss the trends
observed. (4p)

Btoh histogram skewed to the right, with most participants having relatively lower IGM value. However, the variation for altered-smell participants seems to be larger than normal participants. The average of IGM is also higher than altered-smell group. Further more, the altered-smell group seem to have more extremed IGM value than normal group.

