---
title: "Homework 1"
author: "jiaying Ning"
date: "9/23/2020"
output:
  html_document: default
  word_document: default
  pdf_document: default
---


**Problem 1**

load packages
```{r, echo = F, message = FALSE, warning=FALSE}
rm(list = ls())
library(arsenal)
library(dplyr)
library(ggplot2)
```

import data
```{r, echo = F, message = FALSE, warning=FALSE}
setwd("/Users/jiayingning/Documents/Columbia\ Fall2020/P8130/Jiaying\ Ning/P8130_jn2761")
Antibodies<-read.csv("Antibodies.csv")
```

1) Using the entire sample, provide descriptive statistics for all variables of interest.
a) Total sample size for each variable (N); Mean/SD, Median/IQR, Min and Max for continuous
variables; Frequency/Percentages for categorical variables; Number of missing values for each
variable. (8p)
b) Generate a his
togram for the Ig-M values and comment on its shape. (2p)


since the only continuous variable is Antibody_Igm. We will be taking the descriptive statistics measure for this continuous variable.

below are its overall summary and desctiptive value table
```{r Problem 1}
#overall summary
summary(Antibodies)

#print out the class table for each variable in the dataset
t(t(sapply(Antibodies, class)))

#print out the descriptive statistics
descriptivestats = summary(na.omit(Antibodies$Antibody_IgM))
descriptivestats
IQR(na.omit(Antibodies$Antibody_IgM))
sd(na.omit(Antibodies$Antibody_IgM))

```

From the dataset, I observe that only meaningful variable that is numeric is "Antibody_IgM". Variable "AgeCategory","Smell","Gender" are all factor.

For continuous variable:
Antibody_Igm has means of `r mean(na.omit(Antibodies$Antibody_IgM))` with standard deviation of `r sd(na.omit(Antibodies$Antibody_IgM))`, median of `r median(na.omit(Antibodies$Antibody_IgM))` and IQR of `r IQR(na.omit(Antibodies$Antibody_IgM))`.Its min is `r min(na.omit(Antibodies$Antibody_IgM))` and max is `r max(na.omit(Antibodies$Antibody_IgM))`

For categorical variable
the frequency and proportion of agegroup is:
```{r freq age}
#frequency for categorical variable
##frequency of age category
table(Antibodies$AgeCategory)
prop.table(table(Antibodies$AgeCategory))
```
the frequency and proportion of subject's smell status is:
```{r freq smell}
##frequency of smell
table(Antibodies$Smell)
prop.table(table(Antibodies$Smell))
```
the frequency and proportion of smell is:
```{r freq gender}
##frequency of Gender
table(Antibodies$Gender)
prop.table(table(Antibodies$Gender))
```

From the summary table we know that there are 1224 NA values in Antibody_Igm variable. 34 unanswered/other in smell category.

```{r histogram IGM}
ggplot(data=Antibodies,mapping=aes((Antibody_IgM))) + geom_histogram()
```
From the graph showing above, we can see that the distribution is skewing to the right. Most people have low antibody_igm. 

**problem 2**

2) Provide descriptive statistics for all variables, stratified by smell category.
a) Provide descriptive statistics for all the other variables by the two smell categories (normal vs
altered) and present them in a tabular form (see example below). Briefly comment on the
differences observed between the two groups. (6p)
Hint: tableby() in R can easily generate this, but feel free to create your own table in Word.

```{r table}
library(dplyr)
library(furniture)

table1(Antibodies,AgeCategory,Subject,Antibody_IgM,Gender,splitby = "Smell",   row_wise = TRUE)

Antibodies_clear = filter(Antibodies, Smell != "Unanswered/Others")
table1(Antibodies_clear,AgeCategory,Subject,Antibody_IgM,Gender,splitby = "Smell",   row_wise = TRUE)


```
Overall, more people reported having their smell altered than not. This difference in proportion is true among all age group and gender. 


b) Use the Ig-M variable to generate side-by-side histograms and boxplots by smell categories
```{r side by side histogram and boxplot}
ggplot(Antibodies_clear,aes(x=Antibody_IgM,fill=Smell))+geom_histogram()

ggplot(Antibodies_clear, aes(x=Smell, y=Antibody_IgM, fill=Smell)) + 
    geom_boxplot()

```



(normal vs altered). Make sure you label your figures appropriately and briefly discuss the trends
observed. (4p)

Btoh histogram skewed to the right, with most participants having relatively lower IGM value. However, the variation for altered-smell participants seems to be larger than normal participants. The average of IGM is also higher than altered-smell group. Further more, the altered-smell group seem to have more extremed IGM value than normal group.



##problem 2

In this article, author discussed a study done by Professor Shoveller who suggested that cats that eats only one meals a day might feel more satisfied and have healthier body than cats who eats several small meals a day. To present this fact, the author compares this result to another more widely accepted theory, which is that “eating small meals several days is better for cats”. Then the author shortly explained how the study set up and lay out the conclusion of the study.

Summary: The main goal of the study is to examine how feeding frequency impact cats’ mental and physical state. Total of 8 cats with similar age are included in the study, all cats are assigned with the same food. The study used a 2X3 replicated incomplete Latin square design. The study does not specify whether cats are randomly assigned to treatment. Some potential bias might include: 
1.	Even though precise measurement for health and behavior is recorded for each cat, the sample size might still be too small to be generalized to a larger cat population
2.	Different feeder prepares different type of food for their cat, and the quality of food can have strong impact on a cats’ mental and physical state. Therefore, since all cats are assigned with the same canned food in the study, it is not clear whether the food choice can be a potential confounding variable.  

Journal/Newspaper Reference: Leggate, J. (2020, September 24). You might be feeding your cats wrong, according to a new study. Fox News. https://www.foxnews.com/lifestyle/feeding-cats-wrong-new-study
Original Study reference: Camara A, Verbrugghe A, Cargo-Froom C, Hogan K, DeVries TJ, et al. (2020) The daytime feeding frequency affects appetite-regulating hormones, amino acids, physical activity, and respiratory quotient, but not energy expenditure, in adult cats fed regimens for 21 days. PLOS ONE 15(9): e0238522.

