---
title: "Midterm_jn2761"
author: "jiaying Ning"
date: "10/29/2020"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(dplyr)
library(arsenal)
```

## problem 3


1) select a random sample (with replacement) of 16 men from this distribution of 2742 normally distribut4ed observations and calculate the corresponding mean
```{r}
set.seed(8130)
population = rnorm(2742, mean = 174, sd = 7.7)
sample = sample (population, size=16, replace =F)
sample_mean = mean(sample)
```
The mean for the sample of 16 men is `r sample_mean`


2. Repeat the step 1 50 times and plot the distribution of 50 sample means.

```{r}
  l <- list()
result=list()
set.seed(8130)

for (i in 1:50){
l[i] <- list(sample (population, size=16, replace =F))
result[i] = list(mean(l[[i]]))
}
result_df = c(unlist(result))
result_df
```

```{r}
hist(result_df)
```

3. *Comment: the shape of the distribution in part 2 is approximatly normal with most of the value gathered around 174, the true population mean. According to central limited theorem, since our underlying population distribution is normal, even though n <30, then the shape of the samping distribution is also approximately normal.*


## problem 5
```{r}
#Import Data 
vote_df=
   read.csv("./data/VoteNY2018.csv")
```

```{r}
voted_recode=
vote_df %>%
  mutate(
        VOTED=recode(VOTED,`1`="Did not vote",`2`="Voted", `96`="Refused", `97`="Don't know", `98` ="No Response", `99` ="Not in universe"),
        VOTED_SIMPLE = ifelse(VOTED=="Voted","YES","NO"),
         SEX = recode(SEX, `1`= "Male",`2`="Female",`9`="NIU"),
        RACESIMPLE=recode(RACESIMPLE, `1`	= "White", `2`= "Black", `3`= "American Indian or Aleut or Eskimo", `4`= "Asian or Pacific Islander", `5` =	"More than one race", `9`	= "Blank"),
        EDUSIMPLE=recode( EDUSIMPLE, `0` = "No school", `1`= "Some school but no diploma", `2` =	"High school graduate or GED", `3`	= "Some college but no degree", `4`	= "Associate degree",`5`= "Bachelors degree", `6`	= "Masters degree",`7` = "Professional or Doctoral degree")) %>%
  
          relocate(VOTED,VOTED_SIMPLE,AGE,SEX,RACESIMPLE,EDUSIMPLE)


```
```{r}
summary(vote_df$AGE)
summary(vote_df$VOTED)
```

I did not use the coding for AGE since there is no value of 0,99, and 99+ in the current data for age. The minimum data is 18 and the maximum data is 85.

Create a summary table of age, sex, race and education level by voting status (yes – no)
in the 2018 elections. Carefully choose the summary statistics to report based on the variable type.
```{r}
 tble = tableby( VOTED_SIMPLE ~ AGE+SEX+RACESIMPLE+EDUSIMPLE, data=voted_recode)
summary(tble, title = "Descriptive Statistics: voting in 2018",  text=T)

```
3. Using a 5% significance level, evaluate the associations between: voting status and race. You need to state the following: hypotheses, table of expected values, statistical test you chose and why, test statistic, critical value, decision rule and interpretation in the context of the problem.

To test the association between two categorical variable, we will construct a contingency table, calculated the expected cell counts and the chi-square statistics testing for independence. 

- H0: There is no association between voting status and race, voting status and race are independent
- H1: There is association between voting status and race, voting status and race are dependent
```{r}
table2 = table(voted_recode$RACESIMPLE,voted_recode$VOTED_SIMPLE)
```

```{r}

table2
```
Here is the table for expected value
```{r}
#expected_value

library(data.table)
expected_value = data.table(voted= c(883*12/2647,883*158/2647,883*354/2647,883*30/2647,883*2093/2647)
,not_voted=c(1764*12/2647,1764*158/2647,1764*354/2647,1764*30/2647.1764*2093/2647))
expected_value
```

$$ X^2 = sum of( (original value-expected value)^2/expected value )$$
$$df= (row-1) * (columb-1)=4$$

```{r}
chisq.test(table2)
qchisq(0.95,4)
```


- X-squared = 26.082, df = 4, p-value = 3.047e-05
- Critical value: qchisq(0.95,4) = 9.487729
- since Test Value = 26.082 > critical value = 9.487729 We reject the null hypothesis and conclude that the proportions of voting status are different among racial groups at alpha level = 0.05